Further Evidence of Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2022-03-24 src/ms.tex
adding Gaia scripts

\subsection{Ages of stars on the long-period pile-up}
\label{subsec:ages}
The WMB model predicts that hotter stars pile up on the long-period edge at younger ages, producing an age gradient across the edge. In Figure~\ref{fig:ages} we show the \teff-age distributions of stars on the ridge using ages from the CKS and SPOCS \citep{Brewer2018} catalogs, where we use the trapezoidal selection described in \S\ref{subsec:rossby} to select stars on the long-period pile-up. Both catalogs derive spectroscopic parameters from the CKS spectra, but use different pipelines for both the spectroscopic parameters and the isochrone fitting. In both cases, there appears to be an age gradient such that hotter stars are younger on average. However, such a trend is also expected in a sample of main-sequence stars as a natural consequence of the shorter main-sequence lifetimes of hotter, more massive stars. We also find that the dispersion in age is a sensitive function of \teff, with cooler stars on the ridge showing a broader range of ages. This observation could be due to cooler stars populating the ridge for longer periods of time (relative to hotter stars), the lower precision of isochrone ages for cooler stars, or some combination of the two effects.

@@

\begin{figure*}
    \includegraphics[width=\linewidth]{./figures/cmd.pdf}
    \caption{Gaia DR2 color-magnitude diagram. In each panel, a 2-d histogram of all Kepler targets is shown in grayscale, with lighter shades representing a higher number of targets in each cell. In the middle and right panels, the \citet{Santos2021} targets with photometric rotation period detections are overplotted in orange. In the right panel, the CKS long-period pile-up stars identified in this work are shown by blue points and a Gaussian kernel density estimation of the \citet{Hall2021} asteroseismic sample is shown by blue contours.}
    \label{fig:cmd}
\end{figure*}

>>>>

\subsection{Ages of stars on the long-period pile-up}
\label{subsec:ages}
The WMB model predicts that hotter stars pile up on the long-period edge at younger ages, producing an age gradient across the edge. In Figure~\ref{fig:ages} we show the \teff-age distributions of stars on the ridge using ages from the CKS \citep{Fulton2018} and SPOCS \citep{Brewer2018} catalogs, where we use the trapezoidal selection described in \S\ref{subsec:rossby} to select stars on the long-period pile-up. Both catalogs derive spectroscopic parameters from the CKS spectra, but use different pipelines for both the spectroscopic parameters and the isochrone fitting. In both cases, there appears to be an age gradient such that hotter stars are younger on average. However, such a trend is also expected in a sample of main-sequence stars as a natural consequence of the shorter main-sequence lifetimes of hotter, more massive stars. We also find that the dispersion in age is a sensitive function of \teff, with cooler stars on the ridge showing a broader range of ages. This observation could be due to cooler stars populating the ridge for longer periods of time (relative to hotter stars), the lower precision of isochrone ages for cooler stars, or some combination of the two effects.

@@

\begin{figure*}
    \includegraphics[width=\linewidth]{./figures/cmd.pdf}
    \caption{Gaia DR2 color-magnitude diagram. In each panel, a 2-d histogram of all Kepler targets is shown in grayscale, with lighter shades representing a higher number of targets in each cell. The estimated position of the Sun is indicated by the black point, based on the calibration of \citet{Casagrande2018}. In the middle and right panels, the \citet{Santos2021} targets with photometric rotation period detections are overplotted in orange. In the right panel, the CKS long-period pile-up stars identified in this work are shown by blue points and a Gaussian kernel density estimation of the \citet{Hall2021} asteroseismic sample is shown by blue contours.}
    \label{fig:cmd}
\end{figure*}
